---
title: "Introduction and usage"
author: "LG"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction and usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

paste0_vec <- function(...){ paste(..., collapse = ',') }
```

This vignette introduces the package. 

## Data generation

Let's first create some data.    
We will use the package `rsamplestudy` to generate a population and to extract the reference/questioned samples.

Here, we will be generating data from a "Dirichlet-Dirichlet" model.    
Note that this has nothing to do with the models implemented in this package.

```{r pop-generation}
library(rstanBF)
library(dplyr)
library(rsamplestudy)


# set.seed(123)
p <- 5
n <- 100
m <- 10
alpha <- rep(p, p)

list_pop <- rsamplestudy::fun_rdirichlet_population(n, m, p, alpha = alpha)
list_pop

```

### Sampling generation

The package compares two sets of *items* (reference and questioned) and evaluates their support to a pair of hypotheses.

- $H_1$: the questioned items come from the reference source
- $H_2$: the questioned items come from other different sources

We simulate a situation using the function `make_dataset_splits` from the package `rsamplestudy`:

```{r sampling-generation}
k_ref <- 10
k_quest <- 10

# We sample from all sources
sources <- unique(list_pop$df_pop$source)
list_samples_diff <- list_pop$df_pop %>% rsamplestudy::make_dataset_splits(k_ref, k_quest)

source_same <- unique(list_samples_diff$df_reference$source)
list_samples_same <- list_pop$df_pop %>% rsamplestudy::make_dataset_splits(k_ref, k_quest, source_ref = source_same, source_quest = source_same)

list_samples_H <- list(H1 = list_samples_same, H2 = list_samples_diff)
```

These need to be converted to a friendly format (matrices and lists) from `rsamplestudy` output:

```{r data-preparation}
list_data_diff <- stanBF_prepare_rsamplestudy_data(list_pop, list_samples_diff)
list_data_same <- stanBF_prepare_rsamplestudy_data(list_pop, list_samples_same)
```

We see that each list contains the full set of observations, the set of indexes to reference items, and the set of indexes to the questioned items:
```{r}
list_data_diff
```

From now on, we are ready to use the `rstanBF` package.

## Data model

The package supports several models:
```{r}
available_models()
```

## Hyperparameter elicitation

We choose the Dirichlet-Dirichlet model.   

```{r}
model <- 'DirDir'
```

The only hyperparameter which must be provided is $\alpha$, the Dirichlet parameter at the hyperprior level.

It is known since the data has been generated: let's use it, for now.

```{r hyperpriors-manual}
list_hyperpriors <- list(alpha = list_pop$alpha)
```

### Using functions

If one wants to elicit a value for the hyperparameter(s) required by the model, the package supplies the function `stanBF_elicit_hyperpriors`.   
See the documentation for details:

```{r hyperpriors-function}
stanBF_elicit_hyperpriors(list_samples_same$df_background, model, mode_hyperparameter = 'ML', mode_ML = 'naive')
stanBF_elicit_hyperpriors(list_samples_same$df_background, model, mode_hyperparameter = 'ML', mode_ML = 'ML')
```

Compare the estimate with the generating value:

$\alpha = \left( `r paste0_vec(list_pop$alpha)` \right)$

## Sampling

We need to set up the HMC parameters, first:
```{r HMC-params}
n.iter <- 2000
n.burnin <- 200
# n.iter <- 10000
# n.burnin <- 1000

n.chains <- 2 # parallel::detectCores()
n.cores <- 2 # parallel::detectCores()
```

Now we are ready! Let's sample:

```{r HMC-sampling}
stanBF_obj_same <- rstanBF::compute_BF_Stan(list_data_same, model, list_hyperpriors, data_other = NULL, n.iter, n.burnin, n.chains, n.cores, silent = TRUE)
stanBF_obj_diff <- rstanBF::compute_BF_Stan(list_data_diff, model, list_hyperpriors, data_other = NULL, n.iter, n.burnin, n.chains, n.cores, silent = TRUE)
```

```{r HMC-outputs}
stanBF_obj_same
stanBF_obj_diff
```


## Bayes Factor

The objects contain the Bayes Factors:

```{r HMC-BF}
stanBF_obj_same$BF
stanBF_obj_diff$BF
```


## Stan objects

rstanBF objects contain `stanfit` objects (one under $H_1$, one under $H_2$), which are actually the outputs of Stan sampling procedure.

They are available under the `stanfit` property:

```{r, eval = FALSE}
stanBF_obj_same$stanfit$H1
stanBF_obj_same$stanfit$H2
```

One can either further process them using standard tools (`rstan` post-processing functions, `coda`, etc.), or use the supplied functions in the package. 

These functions usually do different things depending on the model (e.g. often only the likelihood).

## Post-processing

What follows must be implemented for each model.   
For now, only `stanBF_turn` models are completely specified.

### Posteriors

Returned objects often implement a `samples` method to extract samples from the interesting posteriors:

```{r, HMC-posteriors}
head(rstanBF::samples(stanBF_obj_same))
head(rstanBF::samples(stanBF_obj_diff))
```

### Plots

Returned objects often implement a `plot_posteriors` method:

```{r, HMC-posteriors-plot, fig.width = 6}
rstanBF::plot_posteriors(stanBF_obj_same, 'theta')
rstanBF::plot_posteriors(stanBF_obj_diff, 'theta')
```


### Prior predictive distributions

Models often implement draws from the prior predictive distributions.

The names of the variables in the Stan source start with `sim_`.   
One can use the `prior_pred` method to extract and format them:

```{r}
df_prior_pred_same <- prior_pred(stanBF_obj_same)
head(df_prior_pred_same, 20)
```

This is interpreted as an example of a dataset, generated according to the model, before having observed any data.

They can be used to assess if the specified hyperparameters are too restrictive or too wide. 

